---
title: "Chapter 8"
output: html_document
---

```{r load_packages}
library(tidyverse)
library(edr)
```

### Listing 8.1. Assigning a character string to `one_word`.

```{r one_word}
one_word <- "hello"
one_word
```

### Listing 8.2. Creating a character vector with three elements.

```{r three_words}
three_words <- c("one", "two", "three")
```

### Listing 8.3. Combining two character vectors with the `c()` function.

```{r words_and_days}
days <- c("monday", "tuesday", "wednesday")

words_and_days <- c(three_words, days)

words_and_days
```

### Listing 8.4. Combining a character vector with a numeric vector.

```{r words_and_numbers}
three_numbers <- c(1, 2, 3)

words_and_numbers <- c(three_words, three_numbers)

words_and_numbers
```

### Listing 8.5. How combining elements of different types results in coercion.

```{r coercion_to_character}
c(7, 1.2, TRUE, 6.4/2, FALSE, "hello")
```

### Listing 8.6. Creating a data frame with two columns.

```{r data_frame_w_factors, paged.print=FALSE}
data_frame <- data.frame(a = three_numbers, b = three_words)
data_frame
```

### Listing 8.7. Using the `stringsAsFactors` option with `data.frame()`.

```{r data_frame_no_factors, paged.print=FALSE}
data_frame_2 <- 
  data.frame(
    a = three_numbers,
    b = three_words,
    stringsAsFactors = FALSE
  )

str(data_frame_2)
```

### Listing 8.x. Using the `toupper()` and `tolower()` functions to transform characters to upper case and lower case.

```{r toupper_tolower}
mixed_cases <- c("One", "two", "Three", "FOUR")

c(
  toupper(mixed_cases[1:2]),
  tolower(mixed_cases[3:4])
)
```



