---
title: "Chapter 6"
output: html_document
---

### Listing 6.1. Loading the tidyverse, lubridate, and rwr packages

```{r load_tidyverse_rwr}
library(tidyverse)
library(rwr)
library(lubridate)
```

### Listing 6.2. Examples of typical date Formats with Different Orderings of Year, Month, and Date

```{r text_dates}
ymd_chr <- "19950615"   # YYYY/MM/DD (ISO 8601)
mdy_chr <- "06/15/1995" # MM/DD/YYYY
dmy_chr <- "15-06-1995" # DD/MM/YYYY
```

### Listing 6.3. Converting text-based dates to R Date objects

```{r to_date_objects}
# Parse each character-based date
date_1 <- ymd(ymd_chr)
date_2 <- mdy(mdy_chr)
date_3 <- dmy(dmy_chr)

# Create a vector of Date objects
dates <- c(date_1, date_2, date_3)

# Print the vector to the console
dates
```

### Listing 6.4. Checking that we have a Date vector

```{r check_class_date}
# Get the class of the `dates` vector
class(dates)
```

### Listing 6.5. Creating a Date object from a date string in ISO notation

```{r as_date_function}
# Use `as_date()` instead of `ymd()`
as_date("19950615")
```

### Listing 6.6. Typical date formats with a time component

```{r text_date_times}  
ymd_hm_chr <- "19950615 0800"    # YYYY/MM/DD HH:MM (ISO 8601)
mdy_hm_chr <- "06/15/1995 08:00" # MM/DD/YYYY HH:MM
dmy_hm_chr <- "15-06-1995 08:00" # DD/MM/YYYY HH:MM
```

### Listing 6.7. Converting text-based date-times to R POSIXct objects

```{r to_datetime_objects}
# Parse each character-based date-time
datetime_1 <- mdy_hm(mdy_hm_chr)
datetime_2 <- dmy_hm(dmy_hm_chr)
datetime_3 <- ymd_hm(ymd_hm_chr)

# Create a vector of POSIXct objects
datetimes <- c(datetime_1, datetime_2, datetime_3)

# Print the vector to the console
datetimes
```

### Listing 6.8. Checking that we have a POSIXct vector

```{r check_class_posixct}
# Get the class of the `datetimes` vector
class(datetimes)
```

### Listing 6.9. The `nycweather` dataset, printed to the console.

```{r print_nycweather, paged.print=FALSE}
nycweather
```

### Listing 6.10. Plotting the dataset (as is) with ggplot

```{r plot_nycweather_asis}
# Make a plot using `nycweather`
ggplot(data = nycweather) +
  geom_point(aes(x = time, y = temp))
```

### Listing 6.11. Plotting the beginning and end of nycweather

```{r plot_nycweather_start_end}
# Take only the first and last 100 rows
nycweather_start_end <-
  dplyr::bind_rows(
    head(nycweather, 100),
    tail(nycweather, 100)
  )

# Make a plot using `nycweather_start_end`
ggplot(data = nycweather_start_end) +
  geom_point(aes(x = time, y = temp))
```

### Listing 6.12. Transforming character times to POSIXct date-times

```{r make_datetime_column, paged.print=FALSE}
nycweather_a <- 
  nycweather %>%
  mutate(time = as_datetime(time))

nycweather_a
```

### Listing 6.13. Using a sample of our dataset for quicker plotting

```{r downsample_for_plotting, paged.print=FALSE}
nycweather_a_samp <-
  nycweather_a %>%
  sample_n(size = 1000) %>%
  filter(!is.na(temp))

nycweather_a_samp
```

### Listing 6.14. Plotting a sample of the revised dataset

```{r plot_with_sampled_data}
ggplot(data = nycweather_a_samp) +
  geom_point(aes(x = time, y = temp))
```

### Listing 6.15. Reducing our dataset to get one measurement per hour

```{r one_per_hour, paged.print=FALSE}
# Get one record per hour by
# grouping by `date` and `hour` and
# keeping only one record per group
nycweather_b <- 
  nycweather_a %>%
  filter(!is.na(temp)) %>%
  mutate(
    date = date(time),
    hour = hour(time)
  ) %>%
  group_by(date, hour) %>%
  slice(1) %>%
  ungroup()

nycweather_b
```

### Listing 6.16. Plotting our hourly temperatures as a line

```{r plot_hourly_data_line}
ggplot(data = nycweather_b) +
  geom_line(aes(x = time, y = temp))
```

### Listing 6.17. Summarizing our data to get daily minimum and maximum temperatures

```{r get_daily_min_max, paged.print=FALSE}
# Summarize daily temperatures into
# `min` and `max` values (grouping
# by `date`)
nycweather_c <-
  nycweather_b %>%
  group_by(date) %>%
  summarize(
    min  = min(temp),
    max  = max(temp)
  )
  
nycweather_c
```

### Listing 6.18. Tidying the table to get one temperature value per row

```{r tidy_table, paged.print=FALSE}
# Use tidyr's gather function to create
# a tidy table daily temperatures by `type`
nycweather_d <- 
  nycweather_c %>%
  tidyr::gather(
    key = "type", value = "temp",
    min, max) %>%
  arrange(date)

nycweather_d
```

### Listing 6.19. Plotting our daily minimum and maximum temperatures

```{r plot_daily_min_max_lines}
ggplot(data = nycweather_d) +
  geom_line(aes(x = date, y = temp, color = type)) +
  scale_color_manual(values = c("red", "blue")) +
  scale_x_date(date_breaks = "2 months", date_minor_breaks = "1 month") + 
  theme_minimal() +
  theme(
    legend.title = element_blank(),
    legend.position = "top",
    legend.justification = "left"
  ) + 
  labs(
    title = "Daily Minimum and Maximum Temperatures",
    subtitle = "2010 Data, Measured at JFK Airport",
    x = "Date",
    y = "Temperature (degrees C)"
  )
```


```{r tidy_up_winniweather, paged.print=FALSE}

winniweather_a <- 
  winniweather %>%
  gather(
    key = hour,
    value = temp,
    starts_with("temp")
  ) %>%
  mutate(hour = case_when(
    hour == "temp00_00" ~ 0L,
    hour == "temp06_00" ~ 6L,
    hour == "temp12_00" ~ 12L,
    hour == "temp18_00" ~ 18L)
  ) %>%
  unite(
    col = "time",
    yearmonth, day,
    sep = "-"
  )

winniweather_a
```

```{r winniweather_date_time, paged.print=FALSE}
winniweather_a %>%
  mutate(date = as_date(time))
```
