---
title: "Chapter 10"
output: html_document
---

```{r load_packages}
library(tidyverse)
library(ggridges)
library(edr)
```


## 10.1. Making Line Graphs

### Listing 10.1. A glimpse at the `employment` dataset.

```{r employment_dataset}
glimpse(employment)
```

### Listing 10.2. A simple line graph.

```{r employ_one_line}
ggplot(employment, aes(x = year, y = unemployed)) +
  geom_line()
```

### Listing 10.3. The simple line graph with some styling and customization applied

```{r employ_one_line_styled}
ggplot(employment, aes(x = year, y = unemployed)) +
  geom_line() +
  scale_x_continuous(breaks = seq(1940, 2010, 10)) +
  scale_y_continuous(
    labels = scales::number_format(suffix = "M", scale = 1e-6, accuracy = 1.0)
  ) + 
  labs(
    title = "Number of Working Age People Unemployed from 1940 to 2010",
    x = NULL, y = "Number Unemployed"
  ) +
  theme_minimal()
```

### Listing 10.4. A line graph with two lines that track closely to each other

```{r employ_recent_two_lines}
employ_recent <- employment %>% slice_tail(n = 10)

ggplot(employ_recent, aes(x = year, y = employed)) +
  geom_line() +
  geom_point(size = 2, shape = 21, fill = "white", stroke = 1) +
  geom_line(aes(x = year, y = nonagriculture), linetype = "dashed", color = "gray") +
  scale_x_continuous(breaks = seq(2000, 2010, 1), minor_breaks = NULL) +
  scale_y_continuous(
    labels = scales::number_format(suffix = "M", scale = 1e-6, accuracy = 1.0)
  ) + 
  labs(
    title = "Total Employed and Those Employed in Anything but Agriculture",
    x = NULL, y = "Number Employed"
  ) +
  theme_minimal()
```

### Listing 10.5. Making a smaller, tidier version of the `employment` dataset with tidyr's `pivot_longer()` function

```{r create_employ_recent_tidy, paged.print=FALSE}
employ_recent_tidy <- 
  employment %>%
  select(year, population, employed, unemployed) %>%
  slice_tail(n = 10) %>%
  pivot_longer(
    cols = population:unemployed,
    names_to = "type",
    values_to = "n"
  ) %>%
  mutate(
    type = factor(type) %>% 
      fct_relevel(c("population", "employed", "unemployed"))
  )

employ_recent_tidy
```

### Listing 10.6. A common pitfall when making line graphs with tidy datasets: a sawtooth pattern

```{r geom_line_sawtooth}
ggplot(employ_recent_tidy, aes(x = year, y = n)) +
  geom_line()
```

### Listing 10.7. Mapping the `linetype` aesthetic to the `type` variable of `employ_recent_tidy` gives us three separate lines in a single `geom_line()` call

```{r geom_line_linetype}
ggplot(employ_recent_tidy) +
  geom_line(aes(x = year, y = n, linetype = type)) +
  scale_x_continuous(breaks = seq(2000, 2010, 1), minor_breaks = NULL) +
  scale_y_continuous(
    labels = scales::number_format(suffix = "M", scale = 1e-6, accuracy = 1.0)
  ) + 
  labs(
    title = "Comparison of Total Population to Employed and Unemployed Citizens",
    x = NULL, y = "Number of Citizens"
  ) +
  theme_minimal()
```

### Listing 10.8. Manually providing linetypes and colors to the three lines is made possible with `scale_linetype_manual()` and `scale_color_manual()`

```{r geom_line_manual_settings}
ggplot(employ_recent_tidy) +
  geom_line(aes(x = year, y = n, linetype = type, color = type)) +
  scale_linetype_manual(values = c("solid", "dashed", "dotted")) +
  scale_color_manual(values = c("black", "steelblue", "red")) +
  scale_x_continuous(breaks = seq(2000, 2010, 1), minor_breaks = NULL) +
  scale_y_continuous(
    labels = scales::number_format(suffix = "M", scale = 1e-6, accuracy = 1.0)
  ) + 
  labs(
    title = "Comparison of Total Population to Employed and Unemployed Citizens",
    x = NULL, y = "Number of Citizens"
  ) +
  theme_minimal()
```

### Listing 10.9. Faceting the three lines is possible with `facet_wrap()` and the legend is probably unnecessary in this arrangement (so it's removed)

```{r geom_line_faceted}
ggplot(employ_recent_tidy) +
  geom_line(aes(x = year, y = n, linetype = type, color = type)) +
  scale_linetype_manual(values = c("solid", "dashed", "dotted")) +
  scale_color_manual(values = c("black", "steelblue", "red")) +
  scale_x_continuous(breaks = seq(2000, 2010, 1), minor_breaks = NULL) +
  scale_y_continuous(
    labels = scales::number_format(suffix = "M", scale = 1e-6, accuracy = 1.0)
  ) + 
  facet_wrap(vars(type), ncol = 1, scales = "free") +
  labs(
    title = "Comparison of Total Population to Employed and Unemployed Citizens",
    x = NULL, y = "Number of Citizens"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

### Listing 10.10. A glimpse at the `rainfall` dataset

```{r rainfall_dataset}
glimpse(rainfall)
```

### Listing 10.11. A simple line plot with a filled area can be made by use of `geom_area()`

```{r rainfall_area}
ggplot(rainfall, aes(x = year, y = r_vancouver)) +
  geom_area()
```

### Listing 10.12. It's more visually satisfying to apply some styles in `geom_area()` like a line color, a fill color, and some transparency.

```{r rainfall_area_color}
ggplot(rainfall, aes(x = year, y = r_vancouver)) +
  geom_area(color = "blue", fill = "lightblue", alpha = 0.4)
```

### Listing 10.13. To complete the look of a line plot with an associated area, a ggplot theme is applied (`theme_bw()`), and useful labels (like a title and a caption) are paid attention to.

```{r rainfall_area_final}
ggplot(rainfall, aes(x = year, y = r_vancouver)) +
  geom_area(color = "blue", fill = "lightblue", alpha = 0.4) +
  labs(
    title = "Annual Total Precipitation for the City of Vancouver",
    caption = "Data source: the rainfall dataset from the edr package.",
    x = NULL, y = "Precipitation, mm"
  ) +
  theme_bw() +
  theme(plot.margin = unit(c(15, 15, 15, 15), "pt"))
```

## 10.2. Working with Bar Plots

### Listing 10.14. Using `geom_col()` instead of `geom_bar(stat = "identity")` to make a simple bar chart of total annual rainfall statistics for Vancouver.

```{r rainfall_bar_1}
ggplot(rainfall, aes(x = year, y = r_vancouver)) +
  geom_col()
```

### Listing 10.15. Making a smaller, tidier version of the `rainfall` dataset with tidyr's `pivot_longer()` function.

```{r rainfall_recent_tidy, paged.print=FALSE}
rainfall_recent_tidy <- 
  rainfall %>%
  slice_head(n = 3) %>%
  pivot_longer(
    cols = starts_with("r"),
    names_to = "city",
    names_prefix = "r_",
    values_to = "precip"
  ) %>%
  mutate(
    city = factor(city) %>% fct_inorder(),
    year = factor(year) %>% fct_inseq()
  )

rainfall_recent_tidy
```

### Listing 10.16. Making a clustered bar chart is now possible with the tidy dataset with `city` and `precip` as new variables. Note the use of `position = "dodge"` to avoid stacking the bars.

```{r rainfall_bar_2}
ggplot(rainfall_recent_tidy, aes(x = year, y = precip, fill = city)) +
  geom_col(position = "dodge")
```

### Listing 10.17. Redefinition of the `x` values as the cities instead of the years, where the `fill` is now mapped to `year`.

```{r rainfall_bar_3}
ggplot(rainfall_recent_tidy, aes(x = city, y = precip, fill = year)) +
  geom_col(position = "dodge")
```

### Listing 10.18. A very presentable version of the previous barplot; there are customizations galore in here (and the extra code is worth it in the end).

```{r rainfall_bar_final}
ggplot(rainfall_recent_tidy, aes(x = city, y = precip, fill = year)) +
  geom_col(position = "dodge", width = 0.7) +
  scale_fill_brewer("Blues") +
  scale_x_discrete(
    labels = c(
      "vancouver" = "Vancouver", "calgary" = "Calgary",
      "kenora" = "Kenora", "toronto" = "Toronto",
      "montreal" = "Montréal", "halifax" = "Halifax",
      "stjohns" = "St. John's"
    )
  ) +
  coord_cartesian(ylim = c(0, 1500)) +
  annotate(
    geom = "text", x = 0.77, y = 1200,
    label = "2017", hjust = 0, angle = 90, size = 3
  ) +
  annotate(
    geom = "text", x = 1.00, y = 1350,
    label = "2018", hjust = 0, angle = 90, size = 3
  ) +
  annotate(
    geom = "text", x = 1.24, y = 960, 
    label = "2019", hjust = 0, angle = 90, size = 3
  ) +
  labs(
    title = "Annual Rainfall Totals",
    subtitle = "Comparison of seven cities for 2017, 2018, and 2019",
    caption = "Data source: the rainfall dataset from the edr package.",
    x = NULL, y = "Precipitation, mm"
  ) +
  theme_minimal() +
  theme(
    panel.grid.major.x = element_blank(),
    legend.position = "none",
    axis.text.x = element_text(vjust = 5)
  )
```

### Listing 10.19. A glimpse at the `german_cities` dataset, which we've used previously

```{r german_cities_dataset}
glimpse(german_cities)
```

### Listing 10.20. An attempt at a stacked bar plot with the `german_cities` dataset is an abject failure (the legend is pretty much all we see).

```{r stacked_bar_plot_1}
german_cities %>%
  ggplot(aes(x = pop_2015, y = state, fill = name)) +
  geom_col()
```

### Listing 10.21. A second attempt at the stacked bar plot now lets us see the plot, and we don't need a legend here anyway.

```{r stacked_bar_plot_2}
german_cities %>%
  ggplot(aes(x = pop_2015, y = state, fill = name)) +
  geom_col(aes(fill = name)) +
  theme(legend.position = "none")
```

### Listing 10.22. Transforming our `german_cities` data is required to get a total population by state (`total_pop`) column and to reorder factor levels (improving the order of bars and the segments within).

```{r german_cities_totals, paged.print=FALSE}
german_cities_totals <- 
  german_cities %>%
  group_by(state) %>%
  arrange(desc(pop_2015)) %>%
  mutate(total_pop = sum(pop_2015)) %>%
  ungroup() %>%
  mutate(
    state = state %>% fct_reorder(total_pop),
    name = name %>% fct_reorder(pop_2015, .desc = TRUE)
  ) %>%
  select(-pop_2011)

german_cities_totals
```

### Listing 10.23. A third attempt at the `german_cities` stacked bar plot: bars are in descending order, cities are stacked by population (increasing, left to right), and the x-axis labels are more readable. 

```{r stacked_bar_plot_3}
german_cities_totals %>%
  ggplot(aes(x = pop_2015, y = state, fill = name)) +
  geom_col() +
  coord_cartesian(xlim = c(0, 9E6)) +
  scale_x_continuous(
    labels = scales::number_format(suffix = "M", scale = 1e-6, accuracy = 1.0)
  ) +
  theme_minimal() +
  theme(
    legend.position = "none", 
    panel.grid.major.y = element_blank()
  )
```

### Listing 10.24. In order to directly label the largest cities in each of the bars, we need a separate dataset of just those dominant cities for `geom_text()`.

```{r german_cities_largest_n, paged.print=FALSE}
german_cities_largest <-
  german_cities_totals %>% 
  group_by(state) %>% 
  arrange(desc(pop_2015)) %>%
  mutate(n = n()) %>%
  slice(1) %>% 
  select(-pop_2015)

german_cities_largest
```

### Listing 10.25. The final stacked bar plot, which gives useful information on large cities within each German state through clearly defined bar segments and helpful annotations.

```{r stacked_bar_plot_final}
german_cities_totals %>%
  ggplot(aes(x = pop_2015, y = state, fill = name)) +
  geom_col(color = "white", size = 0.2) +
  scale_fill_grey(end = 0.7, start = 0.1) +
  geom_text(
    data = german_cities_largest,
    aes(y = state, x = total_pop, label = name),
    size = 3,
    hjust = 0,
    nudge_x = 1e5
  ) +
  geom_text(
    data = german_cities_largest,
    aes(y = state, x = 0, label = paste0("(", n, ")")),
    size = 2.5,
    hjust = 1,
    nudge_x = -1e5
  ) +
  coord_cartesian(xlim = c(0, 9E6)) +
  scale_x_continuous(
    labels = scales::number_format(suffix = "M", scale = 1e-6, accuracy = 1.0),
    breaks = seq(1e6, 9e6, 1e6), minor_breaks = NULL
  ) +
  geom_vline(xintercept = 0, color = "gray50") +
  labs(
    title = "Comparison of Most Populous German Cities Across All States ",
    subtitle = "The largest city in each state is provided at right.\n",
    caption = "Data source: the german_cities dataset from the edr package.",
    x = NULL, y = NULL
  ) +
  theme_minimal() +
  theme(
    legend.position = "none", 
    panel.grid.major.y = element_blank(),
    plot.title.position = "plot",
    plot.caption.position =  "plot"
  )
```

## 10.3. Lollipop Plots and Cleveland Dot Plots

### Listing 10.26. A glimpse at the `nycweather` dataset, which we've used previously.

```{r nycweather_dataset}
glimpse(nycweather)
```

### Listing 10.27. Transforming data in the `nycweather` dataset so that there are monthly summaries of high and low temperatures.

```{r nyc_highlow_temps, paged.print=FALSE}
nyc_highlow_temps <- 
  nycweather %>%
  mutate(
    month = lubridate::month(time, label = TRUE, abbr = FALSE),
    day = lubridate::day(time)
  ) %>%
  group_by(month, day) %>%
  summarize(
    min_temp_d = min(temp, na.rm = TRUE),
    max_temp_d = max(temp, na.rm = TRUE)
  ) %>%
  group_by(month) %>%
  summarize(
    min_temp = min(min_temp_d),
    median_min_temp = median(min_temp_d, na.rm = TRUE),
    median_max_temp = median(max_temp_d, na.rm = TRUE),
    max_temp = max(max_temp_d)
  ) %>%
  pivot_longer(cols = ends_with("temp")) %>%
  mutate(
    month = month %>% fct_rev(),
    name = name %>% fct_relevel(c(
      "min_temp", "median_min_temp",
      "median_max_temp", "max_temp"
    )))

nyc_highlow_temps
```

### Listing 10.28. Creating a basic lollipop plot with data from `nyc_highlow_temps`. 

```{r lollipop_max}
nyc_highlow_temps %>%
  dplyr::filter(name == "max_temp") %>%
  ggplot() +
  geom_segment(aes(x = 0, xend = value, y = month, yend = month), color = "gray75") +
  geom_point(aes(x = value, y = month), color = "red")
```

### Listing 10.29. Creating a more sophisticated lollipop plot using specific colors on the points (blue and red, for below and above zero degrees Celsius).

```{r lollipop_min}
nyc_highlow_temps %>%
  dplyr::filter(name == "min_temp") %>%
  mutate(
    side = if_else(value <= 0, "negative", "positive") %>% 
      as.factor()
  ) %>%
  ggplot() +
  geom_segment(
    aes(x = 0, xend = value, y = month, yend = month),
    color = "gray85", size = 1.5
  ) +
  geom_point(aes(x = value, y = month, color = side), show.legend = FALSE) +
  scale_color_manual(values = c("blue", "red")) +
  coord_cartesian(xlim = c(-10, 20)) +
  labs(
    title = "Monthly Low Temperatures in New York (2010)",
    caption = "\nData source: the nycweather dataset from the edr package.",
    x = "Temperature, ºC", y = NULL
  ) +
  theme_minimal() +
  theme(axis.title.x = element_text(hjust = 1))
```

### Listing 10.30. Creating a basic Cleveland dot plot with `nyc_highlow_temps`.

```{r cleveland_plot_1}
nyc_highlow_temps %>%
  ggplot(aes(x = value, y = month)) +
  geom_line(color = "gray75") +
  geom_point(aes(color = name)) +
  scale_color_manual(values = c("red", "blue", "green", "yellow"))
```

### Listing 10.31. A Cleveland dot plot with more meaningful colors for the points, and, extra touches to make the plot look really nice.

```{r cleveland_plot_final}
nyc_highlow_temps %>%
  mutate(color = case_when(
    name == "min_temp" ~ "blue",
    name == "median_min_temp" ~ "deepskyblue",
    name == "median_max_temp" ~ "coral",
    name == "max_temp" ~ "red"
  )) %>%
  ggplot(aes(x = value, y = month)) +
  geom_line(color = "gray75") +
  geom_point(aes(color = color)) +
  scale_color_identity(guide = "none") +
  scale_x_continuous(
    labels = scales::number_format(suffix = "ºC"),
    limits = c(-10, 40),
    minor_breaks = seq(-10, 40, 1)
  ) +
  labs(
    title = "Monthly Low and High Temperatures in New York (2010)",
    subtitle = "Using daily extreme values and average of daily extremes by month.\n",
    caption = "Data source: the nycweather dataset from the edr package.",
    x = NULL, y = NULL
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom", 
    plot.title.position = "plot",
    plot.caption.position =  "plot",
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_line(color = "gray60", size = 1/5),
    panel.grid.minor.x = element_line(color = "gray80", size = 1/10),
    plot.margin = unit(c(15, 15, 15, 15), "pt")
  )
```

## 10.4. Creating Effective Scatter Plots

### Listing 10.32. A glimpse at the `imdb` dataset.

```{r imdb_dataset}
glimpse(imdb)
```

### Listing 10.33. A scatter plot with 2005-2015 data from the `imdb` dataset.

```{r imdb_plot_linear_y}
imdb %>%
  filter(year %in% 2005:2015) %>%
  ggplot(aes(x = score, y = gross)) +
  geom_point()
```

### Listing 10.34. Transforming the `imdb` dataset for the plot by filtering the years of movies and setting up the `year` variable as a factor.

```{r imdb_filtered, paged.print=FALSE}
imdb_filtered <- 
  imdb %>%
  filter(year %in% 2005:2015) %>%
  mutate(year = as.factor(year) %>% fct_rev())

imdb_filtered
```

### Listing 10.35. A scatter plot using the `imdb_filtered` data; uses gray points according to year of release and transforms y values to a log scale.

```{r imdb_plot_log_y}
imdb_filtered %>%
  ggplot(aes(x = score, y = gross)) +
  geom_point(aes(color = year)) +
  scale_color_grey() +
  scale_y_log10()
```

### Listing 10.36. Getting the median earnings and median rating from `imdb_filtered` to generate dividing lines in the finalized plot.

```{r median_earnings_rating}
median_earnings <- median(imdb_filtered$gross)
median_rating <- median(imdb_filtered$score)

median_earnings
median_rating
```

### Listing 10.37. The final plot of the filtered `imdb` dataset, with customized axes and annotated median value lines.

```{r imdb_final_plot}
imdb_filtered %>%
  ggplot(aes(x = score, y = gross)) +
  geom_point(aes(color = year), alpha = 0.5, position = "jitter") +
  scale_color_grey() +
  scale_y_log10(
    labels = scales::dollar_format(),
    breaks = c(1e2, 1e3, 1e4, 1e5, 1e6, 1e7, 1e8, 1e9) 
  ) +
  scale_x_continuous(
    limits = c(1, 10),
    breaks = 1:10,
    expand = c(0, 0.1), 
  ) +
  geom_hline(
    yintercept = median_earnings,
    linetype = "dashed", color = "forestgreen"
  ) + 
  geom_vline(
    xintercept = median_rating,
    linetype = "dashed", color = "steelblue"
  ) +
  annotate(
    geom = "text", x = 10, y = median_earnings + 1.5E7,
    label = "Median Earnings",
    hjust = 1, size = 2.5
  ) +
  annotate(
    geom = "text", x = median_rating - 0.15, y = 100,
    label = "Median Rating",
    hjust = 0, angle = 90, size = 2.5
  ) +
  labs(
    title = "Comparison of Movies' Gross Earnings Compared to Their IMDB Ratings",
    subtitle = "Over approximately 150 films each year from the 2005-2015 period\n",
    caption = "Data source: the imdb dataset from the edr package.",
    x = "IMDB Rating", y = NULL
  ) +
  theme_bw() +
  theme(
    legend.position = "none",
    plot.title.position = "plot",
    plot.caption.position =  "plot"
  )
```

## 10.5 Plotting Distributions

### Listing 10.38. A glimpse at the `pitchfork` dataset.

```{r pitchfork_dataset}
glimpse(pitchfork)
```

### Listing 10.39. A histogram showing the frequencies of binned scores (0-10) from the `pitchfork` dataset.

```{r pitchfork_histogram}
ggplot(pitchfork) +
  geom_histogram(aes(x = score))
```

### Listing 10.40. Setting a `binwidth` per the recommendation given by the **ggplot** package: using a value of `1` makes sense here.

```{r pitchfork_histogram_binwidth}
ggplot(pitchfork) +
  geom_histogram(aes(x = score), binwidth = 1)
```

### Listing 10.41. Customizing the x axis (to show labels for all score bins) and faceting by year gives some insight on how the score distribution changed with time.

```{r pitchfork_histogram_facets}
ggplot(pitchfork) +
  geom_histogram(aes(x = score), binwidth = 1) +
  scale_x_continuous(breaks = 0:10) +
  facet_wrap(vars(year))
```

### Listing 10.42. Using the year as a categorical variable in a boxplot of Pitchfork album ratings can reveal how ratings tended to change over the years.

```{r pitchfork_boxplot}
pitchfork %>%
  mutate(year = factor(year)) %>%
  ggplot() +
  geom_boxplot(aes(x = year, y = score))
```

### Listing 10.43. A box plot with jittered data points can show us the quantity and distribution of ratings along with the summary statistics.

```{r pitchfork_boxplot_points}
pitchfork %>%
  mutate(year = factor(year)) %>%
  ggplot(aes(x = year, y = score)) +
  geom_boxplot(outlier.shape = NA, color = "steelblue") +
  geom_point(position = "jitter", color = "purple", size = 0.2, alpha = 0.25)
```

### Listing 10.44. A violin plot can be more interpretable than overlaid points on a box plot if the number of data points is overwhelming.

```{r pitchfork_violin_plot}
pitchfork %>%
  mutate(year = factor(year)) %>%
  ggplot() +
  geom_violin(
    aes(x = year, y = score, fill = year),
    draw_quantiles = c(0.25, 0.50, 0.75), 
    show.legend = FALSE
  ) + 
  scale_fill_viridis_d(alpha = 0.5, option = "E")
```

### Listing 10.45. A glimpse at the `dmd` dataset, which we've used previously

```{r dmd_dataset}
glimpse(dmd)
```

### Listing 10.46. Creating a simple density plot, mapping `carats` from the `dmd` dataset to `x`.

```{r dmd_density_plot}
ggplot(dmd, aes(x = carats)) +
  geom_density()
```

### Listing 10.47. The `geom_density()` function has a default bandwidth but modifying it with `adjust` has a strong effect on the plotted density curve. 

```{r dmd_density_bandwidths}
ggplot(dmd, aes(x = carats)) +
  geom_density(adjust = 1, color = "brown", size = 3) +
  geom_density(adjust = 1/2, color = "forestgreen", size = 2) + 
  geom_density(adjust = 1/3, color = "darksalmon", size = 1) +
  geom_density(adjust = 1/4, color = "dodgerblue", size = 0.5)
```

### Listing 10.48. The `dmd` dataset is mutated to add a new column (`dollars_carat`) and to produce factors for better control of ordering facets. 

```{r dmd_mutated, paged.print=FALSE}
dmd_mutated <-
  dmd %>%
  mutate(
    dollars_carat = price / carats,
    color = color %>% fct_rev(),
    cut = cut %>% as.factor(),
    clarity = clarity %>% as.factor()
  )

dmd_mutated
```

### Listing 10.49. With `dmd_mutated`, a set of faceted density plots (through `facet_grid()`) is generated to compare distributions of diamond value by mass.

```{r dmd_density_facets}
ggplot(dmd_mutated) +
  geom_density(
    aes(x = dollars_carat, fill = cut, color = cut),
    alpha = 0.2
  ) + 
  facet_grid(
    rows = vars(color),
    cols = vars(clarity), 
    labeller = label_both
  ) +
  scale_x_continuous(
    labels = scales::dollar_format(suffix = "\n/ct"), 
  ) +
  labs(
    title = "Distributions of USD/Carat Values for Diamonds",
    subtitle = "Uses 2,697 diamonds with varying color, cut, and clarity\n",
    caption = "Data source: the dmd dataset from the edr package.",
    x = NULL, y = NULL
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.text.x = element_text(size = 8)
  )
```

### Listing 10.50. With the functions available in **ggridges**, it's possible to make a compact, ridgeline density plot of IMDB movie ratings over 15 years.

```{r imdb_ggridges}
ggplot(imdb, aes(x = score, y = year, group = year)) +
  geom_density_ridges(
    scale = 3, rel_min_height = 0.01, 
    size = 1, color = "steelblue", fill = "lightblue"
  ) +
  scale_x_continuous(breaks = 0:10) +
  scale_y_reverse(breaks = 2000:2015, expand = c(0, 0)) +
  coord_cartesian(clip = "off", xlim = c(0, 10)) +
  labs(
    title = "Distributions of IMDB Movie Ratings by Year",
    subtitle = "Over approximately 150 films each year from the 2000-2015 period\n",
    caption = "Data source: the imdb dataset from the edr package.",
    x = "IMDB Rating", y = NULL
  ) +
  theme_ridges() +
  theme(
    plot.title.position = "plot",
    plot.caption.position = "plot", 
    axis.text = element_text(size = 10)
  )
```

### Listing 10.51. The comparable ridgeline density plot with 15 years of Pitchfork album reviews makes for a great companion piece to the IMDB plot.

```{r pitchfork_ggridges}
pitchfork %>%
  filter(year <= 2015) %>%
  ggplot(aes(x = score, y = year, group = year)) +
  geom_density_ridges(
    scale = 3, rel_min_height = 0.01, 
    size = 0.5, color = "coral", fill = "#FFE8D2"
  ) +
  scale_x_continuous(breaks = 0:10) +
  scale_y_reverse(breaks = 2000:2015, expand = c(0, 0)) +
  coord_cartesian(clip = "off", xlim = c(0, 10)) +
  labs(
    title = "Distributions of Pitchfork Album Ratings by Year",
    subtitle = "Over approximately 1,000 albums each year from the 2005-2015 period\n",
    caption = "Data source: the pitchfork dataset from the edr package.",
    x = "Pitchfork Rating", y = NULL
  ) +
  theme_ridges() +
  theme(
    plot.title.position = "plot",
    plot.caption.position = "plot", 
    axis.text = element_text(size = 10)
  )
```

### Listing 10.52. The `nycweather` dataset is a natural fit for a ridgeline plot, where temperature distibutions are compared by month in 2010.

```{r nycweather_ggridges}
nycweather %>%
  filter(!is.na(temp)) %>%
  mutate(
    month = lubridate::month(time, label = TRUE, abbr = FALSE),
    tempf = (temp * 9/5) + 32
  ) %>%
  ggplot(aes(x = tempf, y = month, fill = stat(x))) +
  geom_density_ridges_gradient(
    scale = 2, rel_min_height = 0.01, 
    color = "gray50", show.legend = FALSE
  ) +
  scale_fill_viridis_c(option = "E") +
  scale_x_continuous(breaks = seq(10, 100, 10)) +
  labs(
    title = "Distributions of Air Temperatures in New York City by Month",
    subtitle = "Uses nearly 13,000 temperature observations from 2010\n",
    caption = "Data source: the nycweather dataset from the edr package.",
    x = "Temperature, ºF", y = NULL
  ) +
  theme_ridges() +
  theme(
    plot.title.position = "plot",
    plot.caption.position = "plot", 
    axis.text = element_text(size = 10)
  )
```



