---
title: "Chapter 7"
output: html_document
---

```{r load_packages}
library(tidyverse)
```

### Listing 7.1. Creating a numeric vector with the `c()` function

```{r vector_c}
x <- c(1, 2, 3)
```

### Listing 7.2. Combining numeric vectors, again with `c()`

```{r vector_c_vector}
y <- c(x, 4, x)
y
```

### Listing 7.3. Performing math operations and combining the results with `c()`

```{r math_operations}
math_1 <- 5 + 5 * 2
math_2 <- (5 + 5) * 2
math_3 <- ((5 + 5) / (5 - 10)) * 2^2

c(math_1, math_2, math_3)
```

### Listing 7.4. Using math functions and combining the results with `c()`

```{r math_functions}
c(
  sqrt(16),
  log(2.5),
  exp(2.5),
  abs(-5)
)
```

### Listing 7.5. Multiplying a numeric vector by a single number

```{r multiplying_y_by_2}
y * 2
```

### Listing 7.6. Multiplying two numeric vectors of the same length

```{r multiplying_y_by_y}
y * y
```

### Listing 7.7. Multiplying two numeric vectors of different lengths

```{r multiplying_different_length_vectors}
vec_1 <- c(1, 2, 3, 4)
vec_2 <- c(3, 2)

vec_1 * vec_2
```

### Listing 7.8. Subsetting the same vector five different ways

```{r list_of_5_y_subsets}
list(
  a = y[5],
  b = y[1:2],
  c = y[c(5, 1, 5)],
  d = y[c(1, 2, 4:5)],
  e = y[seq(4, 6)]
)
```

### Listing 7.9. Subsetting a vector with a conditional statement

```{r subset_with_conditional}
y[y >= 3]
```

### Listing 7.10. The result of a conditional statement

```{r eval_of_conditional_statement}
y >= 3
```

### Listing 7.11. Creating a simple tibble with the `tibble()` function

```{r example_tibble, paged.print=FALSE}
example_tbl <-
  dplyr::tibble(
    col_1 = 1:10,
    col_2 = letters[1:10],
    col_3 = LETTERS[10:1]
  )

example_tbl
```

### Listing 7.12. Extracting a vector from a tibble with `$`

```{r example_tibble, paged.print=FALSE}
example_tbl$col_1
```

### Listing 7.13 Extracting a vector from a tibble with `[[` and a column name

```{r tibble_extract_2}
example_tbl[["col_1"]]
```

### Listing 7.14 Extracting a vector from a tibble with `[[` and a column index

```{r tibble_extract_3}
example_tbl[[1]]
```

### Listing 7.15 Extracting a vector and performing a check of the values

```{r tibble_extract_check}
example_tbl[[1]] %>%
  {. > 8} %>%
  any()
```

### Listing 7.16 Getting a summary of a tibble

```{r summary_of_tibble}
summary(example_tbl)
```

### Listing 7.17 A function that adds two values

```{r adding_function}
add_2_numbers <- function(x, y) {

  answer <- x + y
  return(answer)
}
```

### Listing 7.18 Using our new function to get the sum of two numbers

```{r calling_add_2_numbers}
add_2_numbers(x = 3, y = 4)
```

### Listing 7.19 Using our new function to get the sum of two numbers

```{r adding_1_number}
add_2_numbers(x = 3)
```

### Listing 7.20 A function with a default value

```{r function_with_a_default}
multiply_three <- function(x, y, z = 3) {

  answer <- x * y * z
  return(answer)
}
```

