---
title: "Chapter 7"
output: html_document
---

```{r load_packages}
library(tidyverse)
library(edr)
```

### Listing 7.1. Creating a numeric vector with the `c()` function

```{r vector_c}
x <- c(1, 2, 3)
```

### Listing 7.2. Combining numeric vectors, again with `c()`

```{r vector_c_vector}
y <- c(x, 4, x)
y
```

### Listing 7.3. Performing math operations and combining the results with `c()`

```{r math_operations}
math_1 <- 5 + 5 * 2
math_2 <- (5 + 5) * 2
math_3 <- ((5 + 5) / (5 - 10)) * 2^2

c(math_1, math_2, math_3)
```

### Listing 7.4. Using math functions and combining the results with `c()`

```{r math_functions}
c(
  sqrt(16),
  log(2.5),
  exp(2.5),
  abs(-5)
)
```

### Listing 7.5. Multiplying a numeric vector by a single number

```{r multiplying_y_by_2}
y * 2
```

### Listing 7.6. Multiplying two numeric vectors of the same length

```{r multiplying_y_by_y}
y * y
```

### Listing 7.7. Multiplying two numeric vectors of different lengths

```{r multiplying_different_length_vectors}
vec_1 <- c(1, 2, 3, 4)
vec_2 <- c(3, 2)

vec_1 * vec_2
```

### Listing 7.8. Subsetting the same vector five different ways

```{r list_of_5_y_subsets}
list(
  a = y[5],
  b = y[1:2],
  c = y[c(5, 1, 5)],
  d = y[c(1, 2, 4:5)],
  e = y[seq(4, 6)]
)
```

### Listing 7.9. Subsetting a vector with a conditional statement

```{r subset_with_conditional}
y[y >= 3]
```

### Listing 7.10. The result of a conditional statement

```{r eval_of_conditional_statement}
y >= 3
```

### Listing 7.11. Creating a simple tibble with the `tibble()` function

```{r example_tibble, paged.print=FALSE}
example_tbl <-
  dplyr::tibble(
    col_1 = 1:10,
    col_2 = letters[1:10],
    col_3 = LETTERS[10:1]
  )

example_tbl
```

### Listing 7.12. Extracting a vector from a tibble with `$`

```{r example_tibble, paged.print=FALSE}
example_tbl$col_1
```

### Listing 7.13 Extracting a vector from a tibble with `[[` and a column name

```{r tibble_extract_2}
example_tbl[["col_1"]]
```

### Listing 7.14 Extracting a vector from a tibble with `[[` and a column index

```{r tibble_extract_3}
example_tbl[[1]]
```

### Listing 7.15 Extracting a vector and performing a check of the values

```{r tibble_extract_check}
example_tbl[[1]] %>%
  {. > 8} %>%
  any()
```

### Listing 7.16 Getting a summary of a tibble

```{r summary_of_tibble}
summary(example_tbl)
```

### Listing 7.17 A function that adds two values

```{r adding_function}
add_2_numbers <- function(x, y) {

  answer <- x + y
  return(answer)
}
```

### Listing 7.18 Using our new function to get the sum of two numbers

```{r calling_add_2_numbers}
add_2_numbers(x = 3, y = 4)
```

### Listing 7.19 Using our new function to get the sum of two numbers

```{r adding_1_number}
add_2_numbers(x = 3)
```

### Listing 7.20 A function with a default value

```{r function_with_a_default}
multiply_three <- function(x, y, z = 3) {

  answer <- x * y * z
  return(answer)
}
```

### Listing 7.21 Code to get sales data on a particular day

```{r get_sales_data_script}
library(tidyverse)
library(edr)

date_str <- "2019-01-15"

sales_day <- 
  sales %>% 
  filter(date == date_str)

order_count <- unique(sales_day$order_id) %>% length()
items_sold <- nrow(sales_day)
total_revenue <- sales_day$price %>% sum()
```

### Listing 7.22 Putting the code into a function

```{r get_sales_data_function}
sales_on_day <- function(day) {
  
  date_str <- day
  
  sales_day <- 
    sales %>% 
    filter(date == date_str)
  
  order_count <- unique(sales_day$order_id) %>% length()
  items_sold <- nrow(sales_day)
  total_revenue <- sales_day$price %>% sum()
  
  sales_day_tbl <-
    tibble(
      date = day,
      order_count = order_count,
      items_sold = items_sold,
      total_revenue = total_revenue
    )
  
  return(sales_day_tbl)
}
```

### Listing 7.23 Calling the new function to see if it works as expected

```{r call_sales_on_day, paged.print=FALSE}
sales_on_day(day = "2019-01-15")
```

### Listing 7.24 Improving the function to summarize all the days

```{r daily_sales_function}
daily_sales <- function(sales_table) {
  
  sales_table %>%
    group_by(date) %>%
    summarize(
      order_count = n_distinct(order_id),
      items_sold = n(),
      total_revenue = sum(price)
    ) %>%
    arrange(desc(date))
}
```

### Listing 7.25 Getting the complete sales summary

```{r call_daily_sales, paged.print=FALSE}
daily_sales(sales_table = sales)
```

### Listing 7.26 Adding an option fetch a summary for specified days

```{r daily_sales_function_2}
daily_sales <- function(sales_table,
                        days = NULL) {
  
  summary_tbl <- 
    sales_table %>%
    group_by(date) %>%
    summarize(
      order_count = n_distinct(order_id),
      items_sold = n(),
      total_revenue = sum(price)
    ) %>%
    arrange(desc(date))
  
  if (!is.null(days)) {
    
    summary_tbl <- 
      summary_tbl %>%
      filter(date %in% days)
  }
  
  summary_tbl
}
```



